---
title: "Example Markdown"
author: "Andrew Moore"
date: "`r Sys.Date()`"
output: pdf_document
---

# Plotting the `midwest` data-set

`midwest` is a data-set contained in the `ggplot2` library. It contains demographics on counties in the midwest, such as *college-achievement*, *income*, and *population density*. I'll be making some visualizations at the county-level eventually, and I'd like to get used to working with this kind of data.

The data can be retrieved after the `ggplot2` package is loaded. The code-chunk below sets up my workspace. We'll start by taking a quick look at the data.

```{r setup, echo = TRUE, message = FALSE, warning = FALSE}
# load the R packages I want to use
library(ggplot2)
library(dplyr)
library(knitr)
library(printr)

# access the dataset stored in R
data(midwest)

# pull out only Michigan's data
midwest <- filter(midwest, state == "MI")

# open up the data, see how it's formatted and look at the first 10 columns
glimpse(midwest[, 1:10])
```

\newpage
The data look tidy, everything seems formatted well; I can probably proceed with some visualizations.
But maybe we should start with a summary table.

```{r}
# create a new data-frame, summarizing data by metro/non-metro counties
my_summary <- midwest %>%
  group_by(inmetro) %>%
  summarize(
    avg_density    = mean(popdensity), 
    med_density    = median(popdensity), 
    avg_perc_white = mean(percwhite)
  )

my_summary %>%
  kable(col.names = c("Metro/Non-Metro", "Average density", "Median density", 
                      "Mean % of white individuals"),
        caption = "Density and % Caucasian, by metropolitan vs non-metropolitan counties",
        digits = 2)
```

Metro counties have a higher level of non-white residents (`r signif(my_summary$avg_perc_white[2], 4)`% white).

